### 2011-12-09 misc/papers/asru11_toolkit_poster/script-lowlevel.tex

\documentclass{article}
\begin{document}

\begin{verbatim}
for n in `get_splits.pl $nj`; do
  featspart[$n]="ark:apply-cmvn --norm-vars=false \
   --utt2spk=ark:$data/split$nj/$n/utt2spk ark:$alidir/$n.cmvn \
   scp:$data/split$nj/$n/feats.scp ark:- | add-deltas ark:- ark:- |"
done
# tree building, graph compilation omitted.
while [ $x -lt $numiters ]; do
   echo "Iteration $x"
   if echo $realign_iters | grep -w $x >/dev/null; then
     echo "Aligning data"
     for n in `get_splits.pl $nj`; do
       $cmd $dir/log/align.$x.$n.log \
         gmm-align-compiled $scale_opts --beam=10 --retry-beam=40 \
           $dir/$x.mdl "ark:gunzip -c $dir/$n.fsts.gz|" "${featspart[$n]}" \
           "ark:|gzip -c >$dir/$n.ali.gz" || touch $dir/.error &
     done
     wait;
     [ -f $dir/.error ] && echo "Alignment error on iteration $x" && exit 1;
   fi
   for n in `get_splits.pl $nj`; do
     $cmd $dir/log/acc.$x.$n.log \
     gmm-acc-stats-ali  $dir/$x.mdl "${featspart[$n]}" \
      "ark,s,cs:gunzip -c $dir/$n.ali.gz|" $dir/$x.$n.acc || touch $dir/.error &
   done
   wait;
   [ -f $dir/.error ] && echo "Accumulation error on iteration $x" && exit 1;
   $cmd $dir/log/update.$x.log \
     gmm-est --write-occs=$dir/$[$x+1].occs --mix-up=$numgauss $dir/$x.mdl \
       "gmm-sum-accs - $dir/$x.*.acc |" $dir/$[$x+1].mdl || exit 1;
   rm $dir/$x.mdl $dir/$x.*.acc
   rm $dir/$x.occs 
   if [[ $x -le $maxiterinc ]]; then 
      numgauss=$[$numgauss+$incgauss];
   fi
   x=$[$x+1];
done
\end{verbatim}
\vspace{0.4in}

\end{document}