### 2011-12-09 misc/papers/asru11_toolkit_poster/script-toplevel.tex

\documentclass{article}
\begin{document}

\begin{verbatim}
## Example of WSJ system building.
local/wsj_data_prep.sh /mnt/speech_data/WSJ?/??-{?,??}.?
local/wsj_prepare_dict.sh
local/wsj_format_data.sh
mfccdir=/mnt/my_storage/kaldi_wsj_mfcc
for x in test_eval92 test_eval93 test_dev93 train_si284; do 
 steps/make_mfcc.sh data/$x exp/make_mfcc/$x $mfccdir 4
done
# skipped some data-subsetting commands here.
# This setup would use GridEngine.
decode_cmd="queue.pl -q queue_name -l ram_free=1200M,mem_free=1200M"
train_cmd="queue.pl -q queue_name -l ram_free=700M,mem_free=700M"

steps/train_mono.sh --num-jobs 10 --cmd "$train_cmd" \
  data/train_si84_2kshort data/lang exp/mono0a
steps/align_deltas.sh --num-jobs 10 --cmd "$train_cmd" \
   data/train_si84_half data/lang exp/mono0a exp/mono0a_ali
steps/train_deltas.sh --num-jobs 10 --cmd "$train_cmd" \
    2000 10000 data/train_si84_half data/lang exp/mono0a_ali exp/tri1
# ... 
\end{verbatim}
\vspace{0.4in}

\end{document}