### 2012-02-10 misc/papers/asru11_toolkit_poster/code-samples.tex

\documentclass{article}
\begin{document}

\begin{verbatim}
void DiagGmm::LogLikelihoods(const VectorBase<BaseFloat> &data,
                             Vector<BaseFloat> *loglikes) const {
  loglikes->Resize(gconsts_.Dim(), kUndefined);
  loglikes->CopyFromVec(gconsts_);
  if (static_cast<int32>(data.Dim()) != Dim()) {
    KALDI_ERR << "DiagGmm::ComponentLogLikelihood, dimension "
        << "mismatch" << (data.Dim()) << " vs. "<< (Dim());
  }
  Vector<BaseFloat> data_sq(data);
  data_sq.ApplyPow(2.0);

  // loglikes +=  means * inv(vars) * data.
  loglikes->AddMatVec(1.0, means_invvars_, kNoTrans, data, 1.0);
  // loglikes += -0.5 * inv(vars) * data_sq.
  loglikes->AddMatVec(-0.5, inv_vars_, kNoTrans, data_sq, 1.0);
}
\end{verbatim}
\vspace{0.4in}

\end{document}