### 2017-05-16 tools/extras/irstlm/doc/LMCompilation.tex

LMs in ARPA, iARPA, and qARPA format can be stored in a compact binary table through the command:

\begin{verbatim}
$> compile-lm train.lm train.blm
\end{verbatim}

\noindent
which generates the binary file {\tt train.blm} that can be quickly loaded in memory.  If the LM
is really very large, {\tt compile-lm} can avoid to create the binary LM directly in memory through the 
option {\tt -memmap=1}, which exploits the {\em Memory Mapping} mechanism in order to work as 
much as possible on disk rather than in RAM. \\

\begin{verbatim}
$> compile-lm train.lm train.blm --memmap=1
\end{verbatim}
\noindent
This option clearly pays a fee  in terms of speed, but  is often the only way to proceed. It is also recommended 
that the hard disk for the LM storage belongs to the computer on which the compilation is performed.

\noindent
Notice that most of the functionalities of {\tt compile-lm} (see below) apply to binary and quantized models.

\noindent
By default, the command uses the directory ``/tmp'' for storing
intermediate results.  For huge LMs, the temporary files can grow
dramatically causing a ``disk full'' system error.  It is possible to
explicitly set the directory used for temporary computation through the
parameter ``--tmpdir''.
\begin{verbatim}
$> compile-lm train.lm train.blm --tmpdir=<mytmpdir>
\end{verbatim}

\subsection{Inverted order of ngrams}
\label{sec:inverted-lm}
For a faster access, the ngrams can be stored in inverted order with the following two commands:
\begin{verbatim}
$> sort-lm.pl -inv -ilm train.lm -olm train.inv.lm
$> compile-lm train.inv.lm train.inv.blm --invert=yes
\end{verbatim}

\paragraph{Warning:} The following pipeline is no more allowed!!

\COMMENT{
or with the following pipeline:
}
\begin{verbatim}
$> cat train.lm | sort-lm.pl -inv | \
   compile-lm /dev/stdin train.inv.blm --invert=yes
\end{verbatim}