### 2017-05-16 tools/extras/irstlm/doc/LMInterpolation.tex

We provide a convenient tool to estimate mixtures of LMs that have been already 
created in one of the available formats.  The tool permits to estimate interpolation
weights through the EM algorithm, to compute the perplexity, and to query the interpolated
LM.

\noindent
Data used in those examples can be found in the directory {\tt example/interpolateLM/},
which represents the relative path for all the parameters of the referred commands.

\noindent
Interpolated LMs are defined by a configuration file in the following format:
\begin{verbatim}
3
0.3 lm-file1
0.3 lm-file2
0.4 lm-file3
\end{verbatim}

\noindent
The first number indicates the number of LMs to be interpolated, then each LM is specified
by its weight and its file (either in ARPA or binary format). Notice that you can interpolate
LMs with different orders\\

\noindent
Given an initial configuration file {\tt lmlist.init} (with arbitrary weights), new weights can be estimated
through Expectation-Maximization on some text sample {\tt test} by running the command:
\begin{verbatim}
$> interpolate-lm lmlist.init --learn=test
\end{verbatim}
\noindent
New weights will be written in the updated configuration file, called by default {\tt lmlist.init.out}.
You can also specify the name of the updated configuration file as follows:

\begin{verbatim}
$> interpolate-lm lmlist.init lmlist.final --learn=test
\end{verbatim}

\noindent
Similarly to {\tt compile-lm}, interpolated LMs can be queried through the option {\tt --score}

\begin{verbatim}
$> interpolate-lm lmlist.final --score=yes < test
\end{verbatim}

\noindent
and can return the perplexity of a given input text (``{\tt --eval text-file}''), optionally  at sentence level  by enabling the option ``{\tt --sentence yes}'',

\begin{verbatim}
$> interpolate-lm lmlist.final --eval=test 
$> interpolate-lm lmlist.final --eval=test --sentence=yes
\end{verbatim}

\bigskip
\noindent
If there are binary LMs in the list,  {\tt interpolate-lm} can avoid to load them in memory through the memory 
mapping option {\tt -memmap=1}.

\noindent
The full list of options is:

\begin{verbatim}
--learn text-file   learn optimal interpolation for text-file
--order n           order of n-grams used in --learn (optional)
--eval text-file    compute perplexity on text-file
--dub dict-size     dictionary upper bound (default 10^7)
--score [yes|no]    compute log-probs of n-grams from stdin
--debug [1-3]       verbose output for --eval option (see compile-lm)
--sentence [yes|no] (compute perplexity at sentence level (identified
                    through the end symbol)
--memmap 1          use memory map to read a binary LM
\end{verbatim}