### 2017-05-16 tools/extras/irstlm/doc/LMPruning.tex

Large LMs files can be pruned in a smart way by means of the command 
{\tt prune-lm} that removes $n$-grams for which resorting to the back-off 
results in a small loss. {\IRSTLM} implements a method similar to the 
Weighted Difference Method described in the paper {\em Scalable Backoff
Language Models} by Seymore and Rosenfeld.

\noindent
The syntax is as follows:
\begin{verbatim}
$> prune-lm --threshold=1e-6,1e-6  train.lm.gz  train.plm
\end{verbatim}
Thresholds for each n-gram level, up from 2-grams, are based on empirical 
evidence. Threshold zero results in no pruning. If less thresholds are specified,
the right most is applied to the higher levels. Hence, in the above example we
could have just specified one threshold, namely {\tt --threshold=1e-6}. 
The effect of pruning is shown in the following messages of {\tt prune-lm}:

\begin{verbatim}1-grams: reading 15059 entries
2-grams: reading 142684 entries
3-grams: reading 293685 entries
done
OOV code is 15058
OOV code is 15058
pruning LM with thresholds: 
 1e-06 1e-06
savetxt: train.plm
save: 15059 1-grams
save: 138252 2-grams
save: 194194 3-grams
\end{verbatim}

\noindent
The saved LM table {\tt train.plm}  contains about 3\% less bigrams, and 34\%  
less trigrams.
Notice that the output of prune-lm is an ARPA LM file, while the input can be 
either an ARPA or binary LM. 
In order to measure the loss in accuracy introduced
by pruning, perplexity of the resulting LM can be computed (see below).

\paragraph{Warning:} the possible quantization should be performed after pruning.

\paragraph{Warning:} 
{\IRSTLM} does not provide a reliable probability for the special
1-gram composed by the ``sentence start symbol'' ({\tt <s>}) , because none
should ever ask for it.  However, this pruning method requires the
computation of the probability of this 1-gram.  Hence, (only) in this case
the probability of this special 1-gram is arbitrarily set to 1.