### 2017-05-16 tools/extras/irstlm/doc/parallelComputation.tex

This package provides facilities to build a gigantic LM in parallel in order to reduce computation time.
The script implementing this feature is based on the {\tt SUN Grid Engine} software\footnote{http://www.sun.com/software/gridware}.

\noindent
To apply the parallel computation run the following script (instead of {\tt build-lm.sh}):

\begin{verbatim}
$> build-lm-qsub.sh -i "gunzip -c train.gz" -n 3  -o train.ilm.gz -k 5
\end{verbatim}
Besides the options of {\tt build-lm.sh}, parameters for the SGE manager can be provided through the following one:

\begin{verbatim}
   -q      parameters for qsub, e.g. "-q <queue>", "-l <resources>"
\end{verbatim}

\noindent
The script performs the same {\em split-and-merge} policy described in Section~\ref{sec:giganticLM}, but some computation is performed in parallel (instead of sequentially) distributing the tasks on several machines.