### 2017-06-08 tools/extras/irstlm/doc/LMFileFormats.tex

{\IRSTLM} supports several types of input and output formats for handling LMs, $n$-gram counts, dictionaries.

%{\IRSTLM} supports three output formats of LMs. These formats have the
%purpose of permitting  the use of LMs by  external programs.

\subsection{File Formats for Dictionary}
The dictionary is the data structure exploited by  {\IRSTLM} to store a set of terms.

{\IRSTLM} saves a dictionary in textual file format consisting of:
\begin{itemize}
\item a header line specifying the most important information about the file itself: the keyword "dictionary", a fixed value 0, and the amount of terms the dictionary contains;
\item a set of terms listed according to either their occurrence or their frequency in the data.
\end{itemize}
Here is an excerpt.
\begin{verbatim}
dictionary 0 7893
<s>
</s>
solemn
ceremony
marks
....
\end{verbatim}

\noindent
Optionally, the occurrence frequencies of each term can be stored as well; in this case the keyword is "DICTIONARY".

\noindent
Here is an excerpt.
\begin{verbatim}
DICTIONARY 0 7893
<s> 5000
</s> 5001
solemn 7
ceremony 59
....
\end{verbatim}

\IMPORTANT{The list order is used by {\IRSTLM} to define the internal codes of the terms.
In the vast majority of cases, it is completely transparent and irrelevant to the user.
Only in very few cases highlighted in this manual, this order is crucial.}

\subsection{File Formats for $n$-gram Table}
The $n$-gram table is the data structure exploited by  {\IRSTLM} to store a set of $n$-grams. {\IRSTLM} stores an $n$-gram table either in textual or binary formats.

\subsubsection{Textual format}
The textual format consists of:
\begin{itemize}
\item a header line specifying the most important information about the file itself: the keyword "nGrAm", the order $n$ of the $n$-grams, the amount of $n$-grams the $n$-gram table contains, and a second keyword representing the table type;
\item a second line reporting the size of the dictionary associated to the $n$-grams;
\item the terms of the dictionary (one term per line) with their frequency;
\item the list of all $n$-grams with their counts.
\end{itemize}
Here is an excerpt.
\begin{verbatim}
NgRaM 3 76857 ngram
7893
<s> 5000
</s> 5001
solemn 7
ceremony 59
...
<s> <s> <s>     2
<s> <s> </s>    1
<s> </s> </s>   1
<s> solemn ceremony     1
<s> a solemn    1
\end{verbatim}

\subsubsection{Binary format}
The binary format is similar, but its main keyword is "{\tt NgRaM}" (different caseing), and the list of $n$-grams is binarized; hence, the last portion of the binary $n$-gram table is not user-readable.

\subsubsection{Google $n$-gram format}
{\IRSTLM} supports the Google $n$-gram format as well both for input and output. This format, always textual, simply consists of the list of all $n$-grams with their counts.
Here is an excerpt.
\begin{verbatim}
<s> <s> <s>     2
<s> <s> </s>    1
<s> </s> </s>   1
<s> solemn ceremony     1
<s> a solemn    1
...
\end{verbatim}

\subsubsection{Table types}
The table type keyword represents the way the $n$-grams are collected and the way they are exploited for further computation:
\begin{itemize}
\item{\tt ngram}: each entry is a standard $n$-gram, i.e. a contiguous sequence of $n$ terms; they are usually used to estimate a standard $n$-gram LM;
\item{\tt co-occ$K$}: each entry is a xxxxxx, where $K$ is XXXXXX;
\item{\tt hm$S$}: each entry is a xxxxxx, where $K$ is XXXXXX;
\end{itemize}

\subsection{File Formats for LM}
{\IRSTLM} handles LM both in textual and binary formats.
It provides facilities to save disk space storing probabilities as quantized values instead of floating  point values, and to reduce access time saving $n$-grams in inverted order.

\subsubsection{Textual Format}
The textual format is the well-known ARPA format introduced in DARPA ASR evaluations  to exchange LMs.
ARPA format  is  supported by most third party LM toolkit, like SRILM and KenLM.

The ARPA format consists of:
\begin{itemize}
\item one block reporting the amount $n$-grams stored for level $m$ of the LM ($m<n$); this block starts with the keyword "{\tt \textbackslash{}data}";
\item one block per level reporting the set of $m$-grams for that level together with their log-probability (first field), and the backoff log-probability (last field); each block starts with the keyword "{\tt \textbackslash{}$m$-grams}", with the correct value for $m$;
\item the keyword  "{\tt \textbackslash{}end\textbackslash{}}" closes the file.
\end{itemize}

\noindent
Here is an excerpt.
\begin{verbatim}
\data\
ngram  1=      7894
ngram  2=     46269
ngram  3=     12188
\1-grams:
-4.79871        <s>     -0.826378
....
-1.20244        <unk>
\2-grams:
-3.29024        <s> <s> -0.221849
...
-0.289359       restructuring of
\3-grams:
-0.397606       <s> <s> <s>
-1.67881        <s> a hong_kong
...
-0.420213       seymf council ,
\end\
\end{verbatim}

\noindent
Empty lines can occur before and after each block.

\noindent
There is no limit to the order $n$ of $n$-grams.

\IMPORTANT{Backoff log-probabilities are not reported if equal to 0; backoff log-probabilities do not exist for the largest order.}

\subsubsection{Quantized Textual Format}
This textual format extends the ARPA textual format including codebooks that quantize 
probabilities and back-off weights of each $n$-gram level.

The quantized ARPA format consists of:
\begin{itemize}
\item a header line specifying the most important information about the file itself: the keyword "qARPA", the order $n$ of the LM, the size of the $n$ codebooks
\item one block reporting the amount $n$-grams stored for level $m$ of the LM ($m<n$); this block starts with the keyword "{\tt \textbackslash{}data}";
\item one block per level reporting first the codebooks for the level and then the set of $m$-grams for that level together with their quantized log probability (first field), and the backoff quantized  log-probability (last field); each block starts with the keyword "{\tt \textbackslash{}$m$-grams}", with the correct value for $m$;
\item the keyword  "{\tt \textbackslash{}end\textbackslash{}}" closes the file.
\end{itemize}

\noindent
Here is an excerpt.
\begin{verbatim}
qARPA 3 256 256 256
\data\
ngram 1= 7894
ngram 2= 46269
ngram 3= 12188
\1-grams:
256
-4.79885 -99
-4.62261 -1.75587
....
0       <s>     53
186     </s>    0
....
\2-grams:
256
-3.79901 -99
-3.62278 -3.01953
...
7       <s>     <s>     255
65      <s>     </s>    255
....
\end\
\end{verbatim}

\subsubsection{Intermediate Textual Format}
This is an {\em intermediate} ARPA format used by {\IRSTLM} for optimizing computation of huge LM. It differs from the ARPA format in two aspects:
\begin{itemize}
\item the header line contains only the keyword {\tt iARPA};
\item the first field of each $n$-gram entry is its smoothed frequency of instead of its log-probability.
\end{itemize}

\COMMENT{
\noindent
Nevertheless, iARPA format is properly managed by the {\tt compile-lm} command
in order to generate a binary version or a standard ARPA version.
}

\subsubsection{Binary Format}
The binary format supported by {\IRSTLM} allows for save disk space and upload the LM quicker.

\noindent 
The binary format consists of:
\begin{itemize}
\item a header line specifying the most important information about the file itself: the keyword "blmt", the order $n$ of the LM, and the amount $n$-grams stored for level $m$ of the LM ($m<n$);
\item a second line reporting the size of the dictionary associated to the $n$-grams;
\item the terms of the dictionary (one term per line) with their frequency, if available;
\item a binary section containing $n$-grams and their probabilities; this portion is not user-readable.
\end{itemize}

\begin{verbatim}
blmt 3       7894      46269      12188
7894
<s> 1
</s> 5001
solemn 7
...
_binary_data_
\end{verbatim}

\subsubsection{Quantized Binary Format}
The quantized binary format stores the quantized version of a LM.

\noindent 
It consists of:
\begin{itemize}
\item a header line specifying the most important information about the file itself: the keyword "Qblmt", the order $n$ of the LM, and the amount $n$-grams stored for level $m$ of the LM ($m<n$);
\item a second line specifying the most important information of the codebooks of each level: the keyword "NumCenters", and  the size of the $n$ codebooks;
\item a third line reporting the size of the dictionary associated to the $n$-grams;
\item the terms of the dictionary (one term per line) with their frequency, if available;
\item a binary section containing the codebooks, the $n$-grams and their quantized probabilities; this portion is not user-readable.
\end{itemize}

\begin{verbatim}
Qblmt 3 7894 46269 12188
NumCenters 256 256 256
7894
<s> 1
</s> 5001
solemn 7
...
_binary_data_
\end{verbatim}

\subsubsection{Inverted Binary Format}
{\IRSTLM} can store the $n$-grams in inverted order to speed up access time. This applies to both standard and quantized binary formats, namely {\tt blmt} or {\tt Qblmt}. The keywords are {\tt blmtI} or {\tt QblmtI}, respectively.