### 2017-06-08 tools/extras/irstlm/doc/LMFiltering.tex

A large LM can be filtered according to a word list through the command:

\begin{verbatim}
$> compile-lm  train.lm filtered.lm --filter=list
\end{verbatim}
The resulting LM will only contain n-grams inside the provided list of words,
with the exception of the 1-gram level, which by default is preserved identical
to the original LM. This behavior can be changed by setting the option 
{\tt --keepunigrams no}.  LM filtering can be useful once very large LMs can
be specialized in advance to work on a particular portion of language.
\noindent
If the original LM is in binary format and is very large, {\tt compile-lm} can avoid to load it in memory,
through the memory mapping option {\tt -memmap=1}.