### 2017-06-08 tools/extras/irstlm/doc/LMInterface.tex

LMs are useful when they can be queried through another application in order to compute 
perplexity scores or n-gram probabilities. {\IRSTLM} provides two possible interfaces: 
\begin{itemize}
\item at the command level, through  {\tt compile-lm}
\item at the c++ library level, mainly through methods of the class {\tt lmtable}
\end{itemize}

\noindent
In the following, we will only focus on the command level interface. Details about
the c++ library interface will be provided in a future version of this manual.

\subsection{Perplexity Computation}
Assume we have estimated and saved the following LM:

\begin{verbatim}
$> tlm -tr=train.www -n=3 -lm=wb -te=test -o=train.lm -ps=no
 n=49984 LP=308057.0419 PP=474.9041687 OVVRate=0.05007602433
\end{verbatim}

\noindent
To compute the perplexity directly from the LM on disk, we can use the command:

\begin{verbatim}
$> compile-lm train.lm  --eval=test
 %% Nw=49984 PP=1064.40 PPwp=589.50 Nbo=38071 Noov=2503 OOV=5.01%
\end{verbatim}
Notice that {\tt PPwp} reports the contribution of OOV words to the perplexity. Each OOV word is indeed penalized by dividing the
LM probability of the {\tt unk} word by  the quantity

\centerline{{\tt DictionaryUpperBound} - {\tt SizeOfDictionary}}

\noindent
The OOV penalty can be modify by changing the {\tt DictionaryUpperBound} with the parameter {\tt --dub} (whose default value is set to $10^7$). \\

\noindent
The perplexity of the pruned LM can be computed with the command:
\begin{verbatim}
$> compile-lm train.plm --eval=test --dub=10000000
%% Nw=49984 PP=1019.69 PPwp=564.73 Nbo=39907 Noov=2503 OOV=5.01%
\end{verbatim}
Interestingly, a slightly better value is obtained which could be explained by the 
fact that pruning has removed many unfrequent trigrams and has redistributed 
their probabilities over more frequent bigrams.

\noindent
Notice that {\tt PPwp} reports the perplexity with a fixed dictionary upper-bound of 10 million words. Indeed:
\begin{verbatim}
$> tlm -tr=train.www -n=3 -lm=wb -te=test -o=train.lm -ps=no -dub=10000000 
n=49984 LP=348396.8632 PP=1064.401254 OVVRate=0.05007602433
\end{verbatim}

\bigskip
\noindent
Again, if the LM is in binary format and is very large, {\tt compile-lm} can avoid to load it in memory,
through the memory mapping option {\tt -memmap=1}.

\bigskip
\noindent
By enabling the option ``{\tt --sentence yes}'', {\tt compile-lm} computes perplexity and related figures (OOV rate, number of backoffs, etc.) for each input sentence. The end of a sentence is identified by a given symbol ({\tt </s>} by default).
\begin{verbatim}
$> compile-lm train.plm --eval=test --dub=10000000 --sentence=yes	
\end{verbatim}
{\small 
\begin{verbatim}
%% sent_Nw=1 sent_PP=23.22 sent_PPwp=0.00 sent_Nbo=0 sent_Noov=0 sent_OOV=0.00%
%% sent_Nw=8 sent_PP=7489.50 sent_PPwp=7356.27 sent_Nbo=7 sent_Noov=2 sent_OOV=25.00%
%% sent_Nw=9 sent_PP=1231.44 sent_PPwp=0.00 sent_Nbo=14 sent_Noov=0 sent_OOV=0.00%
%% sent_Nw=6 sent_PP=27759.10 sent_PPwp=25867.42 sent_Nbo=19 sent_Noov=1 sent_OOV=16.67%
.....
%% sent_Nw=5 sent_PP=378.38 sent_PPwp=0.00 sent_Nbo=39893 sent_Noov=0 sent_OOV=0.00%
%% sent_Nw=15 sent_PP=4300.44 sent_PPwp=2831.89 sent_Nbo=39907 sent_Noov=1 sent_OOV=6.67%
%% Nw=49984 PP=1019.69 PPwp=564.73 Nbo=39907 Noov=2503 OOV=5.01%
\end{verbatim}
}

\bigskip
\noindent
Finally, tracing information with the {\tt --eval }  option are shown by setting 
debug levels from 1 to 4 ({\tt --debug}):
\begin{enumerate}
\item reports the back-off level for each word
\item adds the log-prob 
\item adds the back-off weight
\item check if probabilities sum up to 1.
\end{enumerate}

\subsection{Probability Computations}
Word-by-word log-probabilities  can be computed as well from standard input with the command:
\begin{verbatim}
$> compile-lm train.lm --score=yes < test

> </s>  1 p= NULL
> <s> <unk>     1 p= NULL
> <s> <unk> of  1 p= -3.530047e+00 bo= 2
> <unk> of the  1 p= -1.250668e+00 bo= 1
> of the senate 1 p= -1.170901e+01 bo= 1
> the senate (  1 p= -5.457265e+00 bo= 2
> senate ( <unk>        1 p= -2.166440e+01 bo= 2
....
....
\end{verbatim}

\noindent
the command reports the currently observed n-gram, 
including {\tt\_unk\_} words, a dummy
constant frequency 1, the log-probability of the n-gram, and the 
number of back-offs performed by the LM.

\paragraph{Warning:} All cross-sentence $n$-grams are skipped. The 1-grams with the sentence start symbol are also skipped. In a $n$-grams all words before the sentence start symbol are removed. For $n$-grams, whose size is smaller than the LM order, probability is not computed, but a {\tt NULL} value is returned.