### 2017-06-08 tools/extras/irstlm/doc/LMQuantization.tex

A language model file in ARPA  format, created with the IRST LM toolkit or
with other tools, can be quantized and stored in a compact data structure, 
called language model table.  Quantization can be performed by the command:

\begin{verbatim}
$> quantize-lm  train.lm train.qlm
\end{verbatim}

\noindent
which  generates   the  quantized  version  {\tt train.qlm} that  encodes all probabilities and back-off 
weights in 8 bits. The  output is a  modified ARPA format, called qARPA. Notice that quantized
LMs reduce memory consumptions at the cost of some loss in performance. Moreover, probabilities
of quantized LMs are not supposed to be properly normalized.