### 2017-06-08 tools/extras/irstlm/doc/gettingStarted.tex

After a successful installation, you are ready to use {\IRSTLM}.

\noindent
In this Section, a basic 4-step procedure is given to estimate a LM and to compute its perplexity on a text.
Many changes to this procedure can be done in order to optimize effectiveness and efficiency according to your needs.

\noindent 
Please refer to Section~\ref{sec:commands} to learn more about each IRSTLM commands, and 
to Section~\ref{sec:functions} to get hints about IRSTLM functionalities.

\IMPORTANT{All programs assume that the environment variable {\bf IRSTLM} is correctly set to {\tt /path/to/install/doc}, and that that environment variable {\bf PATH} includes the command directory {\tt /path/to/install/bin}. see above}

\noindent
Data used in the following usage examples can be found in an archive you can download from the official website of {\IRSTLM}.
Most of them are very little, so the reported figures are not reliable.

\subsection{Preparation of Training Data}
In order to estimate a Language Model, you first need to prepare your training corpus. The corpus just consists of a text.
We assume that the text is already preprocessed according to the user needs; this means that lowercasing, uppercasing, tokenization, and any other text transformation has to be performed beforehand with other tools.

\noindent
You can only decide whether you are interested that {\IRSTLM} is aware of sentence boundaries, i.e. where a sentence starts and ends. Otherwise, it considers the corpus as a continuous stream of text, and does not identify sentence splits.

\noindent
The following script adds start and end symbols ({\tt <s>} and {\tt </s>}, respectively) to all lines in your training corpus.
\begin{verbatim}
$> cat train.txt | add-start-end.sh > train.txt.se
\end{verbatim}

\noindent
{\IRSTLM} does not compute probabilities for cross-sentence $n$-grams, i.e. $n$-grams including the pair {\tt </s>  <s>}.

\IMPORTANT{{\IRSTLM} assumes that each line corresponds to a sentence, regardless the presence of punctuation inside or at the end of the line.}
\IMPORTANT{Start and end symbols ({\tt <s>} and {\tt </s>}) should be considered reserved symbols, and used only as sentence boundaries.}

\subsection{Computation of $n$-gram statistics}
\noindent
You can now collect $n$-gram statistics for your training data (3-gram in this example) by running the command:

\begin{verbatim}
$> ngt -i=train.txt.se -n=3 -o=train.www -b=yes
\end{verbatim}

\noindent
The $n$-grams counts are saved in the binary file "train.www".

\subsection{Estimation of the LM}
\noindent
You can now estimate a $n$-gram LM (3-gram LM in this example) smoothed according to the Linear Witten Bell method by running the command:

\begin{verbatim}
$> tlm -tr=train.www -n=3 -lm=LinearWittenBell -obin=train.blm
\end{verbatim}
\noindent
The estimated LM is saved in the binary file "train.blm".

\subsection{Computation of the Perplexity}
\noindent
With the estimated LM, you can now compute the perplexity of any text contained in "test.txt" by running the commands below.

\noindent
To be compliant with the training data actually used to estimate the LM, start and end symbols are added to the text as well.

\begin{verbatim}
$> cat test.txt | add-start-end.sh > test.txt.se
$> compile-lm  train.blm  --eval=test.txt.se
\end{verbatim}

\noindent
which produces the output:
\begin{verbatim}
%% Nw=1009 PP=8547.90 PPwp=6870.51 Nbo=983 Noov=119 OOV=11.79%
\end{verbatim}

\noindent
The output shows the number of  words ({\tt Nw}), the LM perplexity ({\tt PP}), the portion of PP due to the out-of-vocabulary words ({\tt PPwp}), the amount of backoff calls({\tt Nbo}) required for computing PP, the amount of out-of-vocabulary words ({\tt Noov}), and the out-of-vocabulary rate ({\tt OOV}).