### 2017-06-08 tools/extras/irstlm/doc/giganticLM.tex

LM estimation starts with the collection of n-grams and their frequency counters. Then, 
smoothing parameters are estimated for each n-gram level; infrequent n-grams are
possibly pruned and, finally, a LM file is created containing n-grams with probabilities and 
back-off weights.  This procedure can be very demanding in terms of memory and
time if it applied on huge corpora.   We provide here a way to split LM training  into smaller and independent steps, that can be easily distributed among independent processes. The  
procedure relies on a training scripts that makes little use of computer RAM and implements 
the  Witten-Bell smoothing method in an exact way.

\noindent
Before starting, let us create a working directory under {\tt examples}, as many files will be created:

\begin{verbatim}
$> mkdir stat
\end{verbatim}

The script to generate the LM is:

\begin{verbatim}
$> build-lm.sh -i "gunzip -c train.gz" -n 3  -o train.ilm.gz -k 5
\end{verbatim}
where the available options are:

\begin{verbatim}
-i    Input training file e.g. 'gunzip -c train.gz'
-o    Output gzipped LM, e.g. lm.gz
-k    Number of splits (default 5)
-n    Order of language model (default 3)
-t    Directory for temporary files (default ./stat)
-p    Prune singleton n-grams (default false)
-s    Smoothing: witten-bell (default), kneser-ney, improved-kneser-ney 
-b    Include sentence boundary n-grams (optional)
-d    Define subdictionary for n-grams (optional)
-v    Verbose
\end{verbatim}

\noindent
The script splits the estimation procedure into 5 distinct jobs, that are explained in
the following section. There are other options that can be used. We recommend for instance to use pruning of singletons to get smaller LM files. 
Notice that {\tt build-lm.sh} produces a LM file {\tt train.ilm.gz} that is NOT in the final ARPA format, but in
an intermediate format called {\tt iARPA}, that is recognized by the {\tt compile-lm} 
command and by the Moses SMT decoder running with {\IRSTLM}. 
To convert the file into the standard ARPA format you can use the command:

\begin{verbatim}
$> compile-lm train.ilm.gz train.lm --text=yes
\end{verbatim}
this will create the proper ARPA file {\tt lm-final}.
To create a gzipped file you might also use:
\begin{verbatim}
$> compile-lm train.ilm.gz /dev/stdout --text=yes | gzip -c > train.lm.gz
\end{verbatim}

\noindent
In the following sections, we will discuss on LM file formats, on compiling LMs into a 
more compact and efficient binary format, and on querying LMs.

\subsection{Estimating a LM with a Partial Dictionary}

A sub-dictionary can be defined by just taking words occurring more than 5 times ({\tt -pf=5})
and at most the top frequent 5000 words ({\tt -pr=5000}):
\begin{verbatim}
$>dict -i="gunzip -c train.gz" -o=sdict -pr=5000 -pf=5 
\end{verbatim}

\noindent
The LM can be restricted to the defined sub-dictionary with the 
command {\tt build-lm.sh} by using the option {\tt -d}: 
\begin{verbatim}
$> build-lm.sh -i "gunzip -c train.gz" -n 3  -o  train.ilm.gz -k 5 -p -d sdict 
\end{verbatim}

\noindent
Notice that all words outside the sub-dictionary will be mapped into the {\tt <unk>}
class, the probability of which will be directly estimated from the corpus statistics.
A preferable alternative to this approach is to estimate a large LM and then to filter
it according to a list of words (see Filtering a LM).