### 2017-06-08 tools/extras/irstlm/doc/installation.tex

\IMPORTANT{The installation procedure has been tested using the {\tt bash} shell on the following operating systems: Mac OSx 10.6.8 (Snow Leopard), Ubuntu 14.04 LTS (trusty), Scientific Linux release 6.3 (carbon).}

\noindent In order to install {\IRSTLM} on your machine, please perform the following steps.

\subsection{Step 0: Preparation of the Configuration Scripts}
Run the following command to prepare up-to-date configuration scripts.
\begin{verbatim}
$> ./regenerate-makefiles.sh [--force]
\end{verbatim}

\WARNING{Run with the "--force" parameter if you want to recreate all links to the autotools.}

\subsection{Step 1: Configuration of the Compilation}
Run the following command to prepare up-to-date compilation scripts, and to optionally set the installation directory (parameter "{\tt -prefix}".
\begin{verbatim}
$> ./configure [--prefix=/path/to/install] [optional-parameters]
\end{verbatim}

You can set other optional parameters to modify the standard compilation behavior.
\begin{verbatim}
  --enable-doc|--disable-doc
    Enable or Disable (default) creation of documentation
  --enable-trace|--disable-trace
    Enable (default) or Disable trace info at run-time
  --enable-debugging|--disable-debugging
    Enable or Disable (default) debugging info ("-g -O2")
  --enable-profiling|--disable-profiling
    Enable or Disable (default) profiling info
  --enable-caching|--disable-caching
    Enable or Disable (default) internal caches
    to store probs and other info
  --enable-interpolatedsearch|--disable-interpolatedsearch
    Enable or Disable (default) interpolated search for n-grams
  --enable-optimization|--disable-optimization
    Enable or Disable (default) C++ optimization info ("-O3")
\end{verbatim}

\noindent
Run the following command to get more details on the compilation options.
\begin{verbatim}
$> configure --help
\end{verbatim}

\subsection{Step 2: Compilation}

\begin{verbatim}
$> make clean
$> make
\end{verbatim}

\subsection{Step 3: Installation}
\begin{verbatim}
$> make install
\end{verbatim}

\noindent
Libraries and commands are generated,  respectively, under the directories\newline {\tt /path/to/install/lib} and {\tt /path/to/install/bin}.

\noindent
If enabled and PdfLatex is installed, this user manual (in pdf) is generated under the directory\newline {\tt /path/to/install/doc}.

\noindent
Although caching is not enabled by default, it is highly recommended to activate through its compilation flag "{\tt --enable-caching}".
%See Section~\ref{sec:caching} to learn more.

\subsection{Step 4: Environment Settings}
Set the environment variable {\tt IRSTLM} to {\tt /path/to/install}.

\noindent
Include the command directory {\tt /path/to/install/bin} into your environment variable {\tt PATH}.
For instance, you can run the following commands

\begin{verbatim}
$> export IRSTLM=/path/to/install/
$> export PATH=${IRSTLM/bin:${PATH}
\end{verbatim}

\subsection{Step 5: Regression Tests}
If the installation procedure succeeds, you can also run the regression tests to double-check the integrity of the software.
Please go to Section~\ref{sec:regressionTests} to learn hot to run the regression tests.

\noindent Regression tests should be run also in the case of any change made in the source code.