### 2017-06-08 tools/extras/irstlm/doc/introduction.tex

This manual illustrates the functionalities of  the IRST Language  Modeling (LM)  toolkit ({\IRSTLM}). It  should  
put you quickly  in  the condition of:
\begin{itemize}
\item extracting the dictionary from a corpus
\item extracting n-gram statistics from it
\item estimating n-gram LMs using different smoothing criteria
\item saving a LM into several textual and binary file
\item adapting a LM on task-specific data
\item estimating and handling gigantic LMs
\item pruning a LM
\item reducing LM size through quantization
\item querying a LM through a command or script
\end{itemize}

\noindent
{\IRSTLM} features very efficient algorithms and data structures suitable to estimate, store, and access very  large LMs.

\noindent
{\IRSTLM} provides adaptation methods to effectively adapt generic LM to specific task when only little task-related data are available.

\noindent
{\IRSTLM} provides standalone programs for all its functionalities, as well as library for its exploitation in other softwares, like for instance speech recognizers, machine translation decoders, and POS taggers.

\noindent
{\IRSTLM} has been integrated into a popular open source SMT decoder  called {\tt Moses}\footnote{http://www.statmt.org/moses/}, and is compatible with LMs created with other tools, such as the SRILM Tooolkit\footnote{http://www.speech.sri.com/projects/srilm}.

\paragraph{Acknowledgments.}Users of this toolkit  might cite in their publications:
\begin{quote}
M. Federico,  N. Bertoldi,  M. Cettolo, {\em IRSTLM: an Open Source Toolkit for Handling Large Scale Language Models}, Proceedings of Interspeech, Brisbane, Australia, pp. 1618-1621, 2008.
\end{quote}

\noindent
References to introductory material on $n$-gram LMs are given in Appendix~\ref{sec:ReferenceMaterial}.