### 2017-06-08 tools/extras/irstlm/doc/releaseNotes.tex

\IMPORTANT{If present, the index in parentheses refers to the revision number in IRSTLM repository (until 5.60.02) or SourceForge repository (from 5.60.03).}
\subsection{Version 3.2}
\begin{itemize}
\item Quantization of probabilities
\item Efficient run-time data structure for LM querying 
\item Dismissal of MT output format
\end{itemize}

\subsection{Version 4.2}
\begin{itemize}
\item Distinction between open source and internal Irstlm tools
\item More memory efficient versions of binarization and quantization commands
\item Memory mapping of run-time LM
\item Scripts and data structures for the estimation and handling of gigantic LMs 
\item Integration of {\IRSTLM} into Moses Decoder
\end{itemize}

\subsection{Version 5.00}
\begin{itemize}
\item Fixed bug in the documentation 
\item General script {\tt build-lm.sh} for the estimation of large LMs.
\item Management of iARPA file format.
\item Bug fixes
\item Estimation of LM over a partial dictionary.
\end{itemize}

\subsection{Version 5.04}
\begin{itemize}
\item Extended documentation with ShiftBeta smoothing. 
\item Smoothing parameter of ShiftBeta can be set manually.
\item Robust handling for smoothing parameters of ModifiedShiftBeta.
\item Fixed probability checks in TLM.
\item Parallel estimation of gigantic LM through SGE
\item Better management of sub dictionary with build-lm.sh   
\item Minor bug fixes
\end{itemize}

\subsection{Version 5.05}
\begin{itemize}
\item (Optional) computation of OOV penalty in terms of single OOV word instead of OOV class
\item Extended use of OOV penalty to the standard input LM scores of compile-lm. 
\item Minor bug fixes
\end{itemize}

\subsection{Version 5.10}
\begin{itemize}
\item Extended ngt to compute statistics for approximated Kneser-Ney smoothing
\item New implementation of approximated Kneser-Ney smoothing method
\item Minor bug fixes
\item More to be added here ....
\end{itemize}

\subsection{Version 5.20}
\begin{itemize}
\item Improved tracing of back-offs
\item Added command prune-lm  (thanks to Fabio Brugnara)
\item Extended lprob function to supply back-off weight/level information
\item Improved back-off handling of OOV words with quantized LM
\item Added more debug modalities to compile-lm
\item Fixed minor bugs in regression tests
\item Updated documentation
\end{itemize}

\subsection{Version 5.21}
\begin{itemize}
\item Addition of interpolate-lm 
\item Added LM filtering to compile-lm
\item Improved regression tests
\item Integration of interpolated LMs in Moses
\item Extended tests on compilers and platforms
\item Improved documentation with website
\end{itemize}

\subsection{Version 5.22}
\begin{itemize}
\item Use of AutoConf/AutoMake toolkit compilation and installation
\end{itemize}

\subsection{Version 5.30}
\begin{itemize}
\item Support for a safe management of LMs with a total amount of $n$-grams larger than 250 million
\item Use of a new parameter to specify a directory for temporary computation because the default ("/tmp") could be too small
\item Improved a safer method of concatenation of gzipped sub lms
\item Improved management of log files
\end{itemize}

\subsection{Version 5.40}
\begin{itemize}
\item Merging of internal-only tlm code into the public version
\item Updated documentation into the public version
\item Included documentation into the public version
\end{itemize}

\subsection{Version 5.50}

\begin{itemize}
\item {\bf 5.50.01}
\begin{itemize}
\item Binary saving directly with tlm
\item Speed improvement through 
\item Caching of probability and states of $n$-grams in the LM interface
\item Storing of $n$-grams in inverted order
\end{itemize}

\item {\bf 5.50.02}
\begin{itemize}
\item Optional creation of documentation
\item Improved documentation
\item Optional computation of the perplexity at sentence-level
\end{itemize}

\end{itemize}

\subsection{Version 5.60}
\begin{itemize}
\item {\bf 5.60.01}
\begin{itemize}
\item Handling of class/chunk LMs with both compile-lm and interpolate-lm
\item Improved pruning strategy to handle with sentence-start symbols
\item Improved documentation and examples
\end{itemize}
\item {\bf 5.60.02}
\begin{itemize}
\item Code cleanup 
\end{itemize}
\item {\bf 5.60.03 (r404)}
\begin{itemize}
\item Xcode project
\item import from IRSTLM repository (revision 4263)
\end{itemize}

\end{itemize}

\subsection{Version 5.70}
\begin{itemize}
\item {\bf 5.70.01 (r454)}
\begin{itemize}
\item Class-based LM
\item Added improved-kneser-ney smoothing for lm-build-qsub.sh
\item Enabled different singleton pruning policy for each submodel of mixture LM
\item Enabled the possibility to load an existing LM up to a specific level smaller than the actual LM order
\item Code tracing
\item Handling of error codes
\item Handling of long filenames and parameters
\item Improved parallel code compilation
\item Improved documentation and examples
\end{itemize}
\item {\bf 5.70.02 (r469)}
\begin{itemize}
\item Code optimization
\item Common interface for all LM types
\end{itemize}
\end{itemize}

\subsection{Version 5.80}
\begin{itemize}
\item {\bf 5.80.01 (r501)}
\begin{itemize}
\item Facility to {\em beautify} source code
\item Re-activation of filtering on a sub-dictionary
\item Code optimization related to LM dumping
\item Transformation of scripts into Bourne shell scripts 
\end{itemize}
\item {\bf 5.80.03 (r579)}
\begin{itemize}
\item Data selection tool
\item Handling of precision upper- and lower-bounds by means of constants.
\item Improved of Xcode project
\item Improved code compilation
\item Improved documentation
\item Improved handling of help
\item Facility to check whether IRSTLM is compile with or without caching
\end{itemize}
\item {\bf 5.80.05 (r642)}
\begin{itemize}
\item Introduction of {\em namespace irstlm}
\item Code optimization
\item Code compliant with OsX Maverick
\item Support for Redis output format to ngt
\item Support CRC16 algorithm
\item Improved plsa command and regression test
\item Improved handling of tracing
\item Improved handling of help
\item Improved handling of error messages
\end{itemize}
\item {\bf 5.80.06 (r647)}
\begin{itemize}
\item Improved code compilation
\end{itemize}
\item {\bf 5.80.07}
\begin{itemize}
\item Changes to LM smoothing types; removed Good-Turing, added Approximated Modified ShiftBeta, renaming
\item Added an additional pruning method, based on level-dependent pruning frequency
\item Improved code compilation
\item Improved documentation
\item Code cleanup
\item Added support for long names of parameters
\item Improved output format 
\end{itemize}
\item {\bf 5.80.08}
\begin{itemize}
\item Added functionality to score n-grams in isolation
\item Added level-based caches for storing prob, state, and statesize
\item Improved management of tracing assert
\item Improved management of tracing/assert/caching AutoConf compilation flags
\item Improved output format
\item Improved code compilation
\item Code cleanup
\end{itemize}
\end{itemize}

\COMMENT{
\subsection{Version 5.xx}
\begin{itemize}
\item {\bf 5.xx.01}
\begin{itemize}
\item 
\end{itemize}
\end{itemize}
}